
Rcode for inference for means (t-interval or t-test)
===========================================================

 The production of a nationally marketed detergent results in certain
 workers receiving prolonged exposure to the bacillus subtilis enzyme.
 Nineteen workers were tested to determine the effects of these exposures
 on various respiratory functions.  The airflow rate, FEV1, is the ratio of
 a person’s forced expiratory volume to the vital capacity, VC (max. volume
 of air a person can exhale after taking a deep breath).  If the enzyme has
 an effect, it will be to reduce the FEV1/VC ratio. The norm is 0.80 in
 persons with no lung dysfunction.



```{r}
ratio <- c(0.61, 0.7, 0.63, 0.76, 0.67, 0.72, 0.64, 0.82, 0.88, 0.82, 0.78, 
    0.84, 0.83, 0.82, 0.74, 0.85, 0.73, 0.85, 0.87)
```

check summary statistics
```{r}
summary(ratio)
```

```{r}
sd(ratio)
```

Are the data symmetric or approximately normal?
```{r}
qqnorm(ratio)
qqline(ratio)
```

Note that to get a t interval and t test  the same function is used.
Type
````r
?t.test 
````
to check what
options are available.  

A 90% confidence interval can be obtained with
```{r}
t.test(ratio, mu = 0.8, conf.level = 0.9)
```


Save the output in an object called temp.
```{r}
temp <- t.test(ratio, mu = 0.8, conf.level = 0.9)
```

check what you can read from the output
```{r}
names(temp)
```



confidence interval
```{r}
temp$conf.int
```


mean
```{r}
temp$estimate
```



degrees of freedom used in calculating the pvalue or the confidence interval
```{r}
temp$parameter
```

However the research question is whether the ratio is lower than normal. This is a one sided
hypothesis test, namely the alternative is $H_A: \mu < 0.80.$
```{r}
temp.1sided<-t.test(ratio, mu=0.80, alternative="less")
````

The pvalue is `r temp.1sided$p.value`. The data provide evidence that exposure to B. subtilis may reduce the FEV1/VC ratio, but are inconclusive at the 5%  significance level.



The R session information (including the OS info, R version and all
packages used):


```{r}
sessionInfo()
```


```{r}
Sys.time()
```



