Rcode for inference with one sample proportions
===================================================

Simulate 100 coin tosses

```{r }
heads <- rbinom(1, size = 100, prob = .5)
```

Test whether about 50% heads came up in these coin tosses. This uses the normal approximation 
to the binomial distribution with or without the continuity correction

```{r }
prop.test(heads, 100)          # continuity correction TRUE by default
prop.test(heads, 100, correct = FALSE)  # without continuity correction
prop.test(heads, 100,conf.level=0.95)   # confidence level of 0.95 is the default
```

The exact test is based on binomial probabilities:

```{r }
binom.test(heads,100)
```

Save the output in an object named temp.

```{r }
temp<-prop.test(heads, 100) 
```

```names(temp)``` shows us what output is generated and the names of the output objects.


Confidence interval for the proportion of heads

```{r }
temp$conf.int
```

Estimated proportion

```{r }
temp$estimate
```

Pvalue for the hypothesis test $H_0: p=0.5$ versus $H_1: p \ne 0.5$

```{r }
temp$p.value
```

